Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-03 Manuscript/manuscript.Rmd
!!! manuscript edits

Replicate Hi-C data for the GM12878 cell line were downloaded from GEO GSE63525 (Experiments HIC001 to HIC029) [@rao20143d]. The chromosomal coordinates of topologically associating domains throughout the genome were obtained using the arrowhead algorithm from the Juicer toolbox provided by the Aiden Laboratory (https://github.com/aidenlab/juicer) [@durand2016juicebox]. Using the juicer toolbox, we were able to call TADs [??? You called TADs yourself, correct?] on [??? how many] contact matrices at 5 kb, 25 kb, 50 kb, and 100 kb resolutions. Identified TAD boundaries were represented by their location on the linear genome (GRCh37/hg19 human genome assembly), including start and end coordinates for chromosomes 1 through 22 (Figure 18 A). The start and end coordinates of all hierarchical TADs were concatenated, sorted, and unique TAD boundary points were obtained (Figure 18 B).

@@

To establish the data for modeling, the genome was binned into equally sized intervals defined by the data resolution (Figure 18 C). For example, for a 5kb resolution contact matrix, the genome was binned into 5kb intervals. To accommodate for the uncertainty in exact boundary location, the boundary points were flanked on both sized by regions of size of the resolution of the corresponding contact matrix. For example, for the domain data obtained from a 5kb resolution contact matrix, each uniquely identified TAD boundary was flanked on either side by 5kb for a total width around each boundary point of 10kb. A genomic bin was labeled as having a TAD boundary within it (Y=1) if it overlapped with a particular flanked boundary. Otherwise, the bin was labeled as not containing a boundary (Y=0).

@@

Genomic annotation data (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Table 1). The annotations consisted of a variety of different chromatin states, histone marks, and DNase I hypersensitive sites.

@@

An ensemble of models was established for the purposes of systematically comparing the performances of different parameters in the framework. Each system in the framework is assumed to be composed of a model constructed from a particular data resolution, with a particular predictor type as the feature space, as follows.

The data set used for modeling was composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the majority and minority sets. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. The training set was then used to perform all downstream modeling (Figure 3). [??? Where is the resampling techniques come?]

### VARIABLE REDUCTION [??? To be modified for using LASSO only]

We evaluated the use of two variable reduction algorithms, the LASSO and Elastic-Net regularizations. Each algorithm was performed prior to re-sampling. For the LASSO, the $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10. The additional $\alpha$ term for the Elastic-Net was tuned over 10 equa-distant values from 0 to 1.

### RE-SAMPLING TECHNIQUES

We then evaluated three re-sampling techniques (in addition to no re-sampling) used to create balanced classes. They are detailed below:

@@

>>>>

Replicate Hi-C data for the GM12878 cell line were downloaded from GEO GSE63525 (Experiments HIC001 to HIC029) [@rao20143d]. The chromosomal coordinates of topologically associating domains throughout the genome were obtained using the arrowhead algorithm from the Juicer toolbox provided by the Aiden Laboratory (https://github.com/aidenlab/juicer) [@durand2016juicebox]. Using the juicer toolbox, we were able to call TADs [??? You called TADs yourself, correct? !!!Yes, using the arrowhead algoirthm from the juicer toolbox] on [??? how many; !!! the first 29 hic matrices (HIC001 to HIC029); should I provide more detail?] contact matrices at 5 kb, 25 kb, 50 kb, and 100 kb resolutions. Identified TAD boundaries were represented by their location on the linear genome (GRCh37/hg19 human genome assembly), including start and end coordinates for chromosomes 1 through 22. The start and end coordinates of all hierarchical TADs were concatenated, sorted, and unique TAD boundary points were obtained.

@@

To establish the data for modeling, the genome was binned into equally sized intervals defined by the data resolution (Figure 1). For example, for a 5kb resolution contact matrix, the genome was binned into 5kb intervals. To accommodate for the uncertainty in exact boundary location, the boundary points were flanked on both sized by regions of size of the resolution of the corresponding contact matrix. For example, for the domain data obtained from a 5kb resolution contact matrix, each uniquely identified TAD boundary was flanked on either side by 5kb for a total width around each boundary point of 10kb. A genomic bin was labeled as having a TAD boundary within it (Y=1) if it overlapped with a particular flanked boundary. Otherwise, the bin was labeled as not containing a boundary (Y=0).

@@

Genomic annotation data (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Supplementary Table 1). The annotations consisted of a variety of different chromatin states, histone marks, and DNase I hypersensitive sites.

@@

An ensemble of models was established for the purposes of systematically comparing the performances of different parameters in the framework. Each system in the framework is assumed to be composed of a model constructed from a particular data resolution, with a particular predictor type for the feature space (Figure 3).

The data sets used for modeling was composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the majority and minority sets. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For each training set, an $l_{1}$ norm penalized regularization (LASSO) was performed. The $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10.

[??? To be modified for using LASSO only; !!!removed variable reduction paragraph and combined it above]

[??? Where is the resampling techniques come?; !!!mentioned it below]
### RE-SAMPLING TECHNIQUES

We then evaluated three re-sampling techniques (in addition to no re-sampling) on the reduced training set, in order to evaluate the problem of imbalanced classes. They are detailed below:

@@

## Figure 1: Model Construction

Diagram of the model construction used for downstream analysis. The linear genome was binned according to the resolution of the respective HiC experiment (either 5kb, 25kb, 50kb, or 100kb intervals). TAD boundaries were then flanked by 1-unit on either side of the boundary point. The unit flanking was indicative of the resolution that the domain data was obtained from (i.e. for 5kb resolution, 1 unit represents 5kb for a total flanking region of 10kb). The response vector Y used for classification was determined by whether or not a genomic bin overlapped with a flanked region. The positional coordinates of each functional genomic element, obtained from ENCODE, were then used to define the feature space of the models.
